## Technical Description

Data for this project comes from the Electric Vehicle Population Data, a publicly available dataset from the Washington State Department of Licensing. This is a huge set of data about the EV registered with Washington State and has the following information: 

Vehicle - make and model
Type - Battery Electric Vehicle or Plug-in Hybrid Electric Vehicle
Electric Range 
Base MSRP  
Geography - County
City of registration 

The data set is a treasure trove of information on the EV diffusion processes in one of the leading regions of the United States.

Data Collection and Format
This information has been collected from lists of vehicles that the Washington State Department of Licensing maintains. Data is administratively collected through registration and updating vehicle lists. The dataset is in CSV format and presents 210,165 records and 17 variables. The presented dataset combines numeric, categorical, and geographic variables.

Update Frequency
While frequency is not directly stated, all vehicle registration datasets do naturally update with time, as well as with the passing of ownership and other changes.


## Missing value analysis
Known Issues and Limitations
The following are issues that were noted during exploration:

Missing Values: 
Some missing values do exist in a few columns, like Legislative District, Vehicle Location, and Electric Range. This reduces the scope when analyzing some of these variables.

Geographical scope:
The given dataset doesn't enable the view on nationwide or global views of how electric vehicle adoption is going on; it's restricted to Washington State only. 

Outdated records: 
Records may not get updated often, and recent registrations or changes may not be reflected. 

Data Cleaning: 
Categorical values or geographic identifiers may have inconsistencies that will need cleaning. 

Data Import Plan:
This is freely downloadable from the Data.gov catalog, in CSV format. First, this data was imported into this project using Python, where preliminary cleaning and some exploration had been done.  Later, it would be loaded into R for analysis and visualization. This dataset provides a robust foundation for analyzing EV adoption and identifying trends, despite its geographic limitations and minor data quality issues.


```{r}

#install.packages("magick")
library(magick)

# Load the image
img <- image_read("~/Desktop/edavfinalproj1.png")

# Plot the image
plot(img)



```
Bar Graph: A graph representing the percent of missed values within the columns; the worst hit is found in the legislative District. The above bar plot quantitatively views the missing data in terms of their percentages. The highest rate of missing value is in the Legislative District column which is about 0.21%, hence the most affected variable in the dataset. Other columns include Vehicle Location, Electric Range, and Base MSRP. These all contain very few missing values, under 0.01%, and are relatively complete, probably without a huge effect on the analysis. Overall, this data is pretty complete; there are only a few columns in the entire set with missing information.

```{r}

# Load the image
img <- image_read("~/Desktop/edavfinalproj2.png")

# Plot the image
image_scale(img,"80%") 
#plot(img)

```

The following heatmap shows the pattern of missing values across the dataset. Each vertical band here represents a column, with highlighted areas showing missing values. Most of the rows are fully filled, and the rest hold few missing values. There is no indication that the missing values clustered in rows, which therefore are most likely to be random and not related to some particular sub-set of the data. The columns most affected-only showing continuous gaps-include the Legislative District. This heat map confirms that the missing data points are not pervasive and affect only a small subset of the data.


